---
title: "Midwestern Presidential Votes"
author: "Jason Heppler"
date: "1/30/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE}
library(leaflet)
library(rio)
library(tidyr)
library(scales)
library(tigris)
library(dplyr)
library(htmlwidgets)

###--------------------------------------------------
### Merge census county-level dataset with map data
###--------------------------------------------------

library(readr)
# Read in data and join to shapefile
datafile <- "data/2016_US_County_Level_Presidential_Results.csv"
flippedCounties <- "data/US_County_Level_Presidential_Results_08-16.csv"
election <- read_csv(datafile)
flipped <- read_csv(flippedCounties)

# Subset county shapefiles to midwestern states and 
# generate a combined FIPS to match against 
# the datafile FIPS.
shp_data <- counties(state = c("MN","SD", "ND", "NE", "WI", "IA", "IL", "OH", "MI", "IN", "MO", "KS"), cb = TRUE, resolution = "20m")
shp_data@data <- shp_data@data %>%
                    mutate(combined_fips = paste(STATEFP, COUNTYFP, sep=""))

# Merge data
library(raster)
midwestMap <-  merge(shp_data, election, by.x = "combined_fips", by.y = "combined_fips")
midwestFlippedData <- merge(midwestMap, flipped, by.x = "combined_fips", by.y = "fips_code")

###--------------------------------------------------
### Data analysis and statistics
###--------------------------------------------------

# Determine which counties flipped in the 2016 election
## thinking through this: if 2012 dem < 2012 gop = flipped
## if 2012 dem >= 2012 dem = incumbent
## if 2012 gop >= 2012 gop = incumbent
## if 2012 gop < 2016 gop = flipped
## 1 = dem flip
## 2 = gop flip
## 3 = incumbent
# TODO: fix this function
midwestFlippedData@data$flipped_counties <- 
  ifelse(midwestFlippedData@data$dem_2016 > midwestFlippedData@data$gop_2016 &
        midwestFlippedData@data$dem_2008 < midwestFlippedData@data$gop_2008, 1,
        ifelse(midwestFlippedData@data$dem_2016 < midwestFlippedData@data$gop_2016 &
               midwestFlippedData@data$dem_2008 > midwestFlippedData@data$gop_2008, 2, 0))
        
               ifelse(midwestFlippedData@data$dem_2016 > midwestFlippedData@data$gop_2016 &
                        midwestFlippedData@data$dem_2008 > midwestFlippedData@data$gop_2008, 0,
                      ifelse(midwestFlippedData@data$dem_2016 < midwestFlippedData@data$gop_2016 &
                               midwestFlippedData@data$dem_2008 < midwestFlippedData@data$gop_2012, 0, 1))))


midwestFlippedData@data$flipped_counties <-
  if (midwestFlippedData@data$dem_2016 > midwestFlippedData@data$gop_2016 &
      midwestFlippedData@data$dem_2012 < midwestFlippedData@data$gop_2012) {
    as.integer(1)
  } else if (midwestFlippedData@data$dem_2016 < midwestFlippedData@data$gop_2016 &
                 midwestFlippedData@data$dem_2012 > midwestFlippedData@data$gop_2012) {
    as.integer(2)
  } else if (midwestFlippedData@data$dem_2016 > midwestFlippedData@data$gop_2016 &
                 midwestFlippedData@data$dem_2012 > midwestFlippedData@data$gop_2012) {
    as.integer(0)
  } else if (midwestFlippedData@data$dem_2016 < midwestFlippedData@data$gop_2016 &
                 midwestFlippedData@data$dem_2012 < midwestFlippedData@data$gop_2012) {
    as.integer(0)
  }

# Factor this flip variable to enable graphing
midwestFlippedData@data$flipped_counties <- as.factor(midwestFlippedData@data$flipped_counties)

# Find min and max percentages for the palette
minPct <- min(c(midwestMap@data$per_dem, midwestMap@data$per_gop))
maxPct <- max(c(midwestMap@data$per_dem, midwestMap@data$per_gop))

###--------------------------------------------------
### Make maps
###--------------------------------------------------

midwestPalette <- colorBin(palette = "Blues", domain = c(minPct, maxPct), bins = c(0, .3, .4, .5, .6, .7, .8, .9, 1), pretty = FALSE)
midwestFlippedPalette <- colorBin(palette = "RdBu", domain = c(0, 1, 2), bins = 3, pretty = FALSE)

midwestPopup <- paste0("<b>", "County: ", "</b>", midwestMap@data$county_name, "<br>",
                       "<b>", "GOP votes (2016): ", "</b>", midwestMap@data$votes_gop, "<br>",
                       "<b>", "Dem votes (2016): ", "</b>", midwestMap@data$votes_dem, "<br>",
                       "<b>", "Difference (2016): ", "</b>", midwestMap@data$diff, "<br>")

# County level votes
midwestResults <- leaflet(shp_data) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(stroke = TRUE,
              smoothFactor = 0.2,
              weight = 1,
              fillOpacity = 0.6,
              color = ~midwestPalette(midwestMap@data$per_dem),
              popup = midwestPopup,
              group = "Democratic Vote"
             ) %>%
  addPolygons(stroke = TRUE,
              smoothFactor = 0.2,
              weight = 1,
              fillOpacity = 0.6,
              color = ~midwestPalette(midwestMap@data$per_gop),
              popup = midwestPopup,
              group = "Republican Vote"
              ) %>%
   addPolygons(stroke = TRUE,
              smoothFactor = 0.2,
              weight = 1,
              fillOpacity = 0.6,
              color = ~midwestPalette(midwestFlippedData@data$flipped_counties),
              popup = midwestPopup,
              group = "Flipped Counties"
              ) %>%
  addLegend(position="bottomright", title = "Flipped Counties", colors=c('#2166ac', '#f5f5f5', '#b2182b'), labels=c("Dem", "Incumbent", "Repub"))  %>%
  addLegend("bottomright", pal = midwestPalette, values = ~midwestMap@data$per_dem, title = "Results",
             labFormat = labelFormat(suffix = '%', between = '% - ',
                                     transform = function(x) 100 * x)
           ) %>%
  addLayersControl(
    baseGroups=c("Democratic Vote", "Republican Vote", "Flipped Counties"),
    position = "bottomleft",
    options = layersControlOptions(collapsed = FALSE)
  )

# Results
midwestResults
```

Farm-to-table hexagon tattooed, pitchfork la croix williamsburg vexillologist ugh neutra distillery. Poke blog activated charcoal selfies, fam mumblecore kale chips trust fund bitters stumptown dreamcatcher. Lo-fi tousled wayfarers vegan, readymade fanny pack organic blog. You probably haven't heard of them jianbing post-ironic cray. Drinking vinegar food truck activated charcoal poutine, affogato williamsburg mixtape health goth swag gentrify chartreuse bicycle rights. Offal pug fixie wayfarers flexitarian. Sartorial deep v coloring book blue bottle, gentrify franzen ennui disrupt keytar organic.


```{r}
library(ggplot2)
library(ggalt)

election.sub <- election %>%
  filter(state_abbr == "NE")

gg <- ggplot()
gg <- gg + geom_dumbbell(data=election.sub,
                        aes(x = per_dem, xend = per_gop,
                            y = county_name),
                        color="#b2b2b2",
                        size = 0.5,
                        point.colour.l = "#0e668b",
                        point.colour.r = "#8b100e")
gg <- gg + labs(x = NULL, y = NULL)
gg <- gg + theme_bw()
gg <- gg + theme(axis.title = element_text(family = "Lato Black", color="#666666", face="bold", size=6)) 
gg <- gg + theme(plot.background=element_rect(fill="#f7f7f7"))
gg <- gg + theme(panel.background=element_rect(fill="#f7f7f7"))
gg <- gg + theme(panel.grid.minor=element_blank())
gg <- gg + theme(panel.grid.major.y=element_blank())
gg <- gg + theme(panel.grid.major.x=element_line())
gg <- gg+ geom_vline(xintercept = .50)
gg <- gg+ geom_vline(xintercept = .20, linetype="dotted", colour="lightgray")
gg <- gg+ geom_vline(xintercept = .40, linetype="dotted", colour="lightgray")
gg <- gg+ geom_vline(xintercept = .60, linetype="dotted", colour="lightgray")
gg <- gg+ geom_vline(xintercept = .80, linetype="dotted", colour="lightgray")
gg <- gg + theme(axis.ticks=element_blank())
gg <- gg + theme(plot.title=element_text(face="bold", family="Lato Regular", size=14))
gg <- gg + theme(plot.subtitle=element_text(family="Lato Regular", size=12))
gg <- gg + theme(plot.caption=element_text(family="Lato Regular", size=10, color="gray", margin=margin(t=10, r=80)))
gg <- gg + labs(title = "Support for Clinton and Trump, Nebraska",
                subtitle = "Percentage of votes for Clinton and Trump in 2016",
                caption="Jason A. Heppler")
gg <- gg + theme(panel.border=element_blank())
gg <- gg + theme(axis.text.y = element_text(family="Lato Regular", size=8))
gg <- gg + theme(legend.position="none")

```